Parsing Postfix log files

### 2009-02-22 doc/logparser-acronyms.tex
The beginning of the architecture rules section.

\newglossaryentry{NLP}{
    type=\acronymtype,
    name={NLP},
    description={Natural Language Processing},
    first={Natural Language Processing (NLP)}
}
\newglossaryentry{NLP glossary}{
    name={Natural Language Processing},
    description={
        Natural Language Processing~\cite{Wikipedia-nlp} attempts to
        increase our understanding of the languages normally used by humans
        (e.g.\ English, Spanish, Japanese), with the goal of writing
        programs that can understand human languages.
    }
}

>>>>

\newglossaryentry{NLP}{
    type=\acronymtype,
    name={NLP},
    description={Natural Language Processing},
    first={Natural Language Processing (NLP)}
}
\newglossaryentry{NLP glossary}{
    name={Natural Language Processing},
    description={
        Natural Language Processing~\cite{Wikipedia-nlp} is an area of
        research that attempts to increase our understanding of the
        languages normally used by humans (e.g.\ English, Spanish,
        Japanese), with the goal of writing programs that can understand
        human languages.
    }
}

\newglossaryentry{FA}{
    type=\acronymtype,
    name={FA},
    description={Finite Automata},
    first={Finite Automata (FA)}
}
\newglossaryentry{FA glossary}{
    name={Finite Automata},
    description={
        Finite automata are computing devices that accept or recognize
        regular languages.  They lack any form of storage, so while they
        can recognise languages such as ${(ab)}^{*}$ or
        $a^{*}{(b^{*}|c)}b$, they cannot count and so cannot recognise
        languages such as $a^{n}b^{n}$.
    }
}

\newglossaryentry{CFL}{
    type=\acronymtype,
    name={CFL},
    description={Context Free Language},
    first={Context Free Language (CFL)}
}
\newglossaryentry{CFL glossary}{
    name={Context Free Language},
    description={
        XXX WRITE THIS DESCRIPTION
    }
}

\newglossaryentry{PDA}{
    type=\acronymtype,
    name={PDA},
    description={Push-Down Automata},
    first={Push-Down Automata (PDA)}
}
\newglossaryentry{PDA glossary}{
    name={Push-Down Automata},
    description={
        XXX WRITE THIS DESCRIPTION
    }
}

%\newglossaryentry{}{
%    type=\acronymtype,
%    name={},
%    description={},
%    first={}
%}
%\newglossaryentry{}{
%    name={},
%    description={
%    }
%}

### 2009-02-22 doc/logparser-architecture.tex
The beginning of the architecture rules section.

@@

Rule have certain characteristics that may help in understanding the
parser:

@@

    \item Rules are annotated with the name of a Postfix program, and will
        only be used when parsing log lines produced by that
        program.\footnote{There are also generic rules which are used when
        parsing log lines produced by any Postfix program, but only if
        there are also rules specific to that program, and those rules have
        already have been tried and have failed on the current log line.}
        Any given rule will only be used to parse a subset of the log
        lines, and any given log line will only have a subset of the rules
        tried against it.

@@

\subsection{Overlapping Rules}

\label{overlapping rules in architecture}

XXX IS THERE ANY PREVIOUS RESEARCH IN THIS AREA\@?

XXX DFA COMPARISON\@.

The parser does not try to detect overlapping rules; that responsibility is
left to the author of the rules.  Unintentionally overlapping rules lead to
inconsistent parsing and data extraction because the order in which rules
are tried against each line may change between log files, and the first
matching rule wins.  Overlapping rules are frequently a requirement,
allowing a more specific rule to match some log lines and a more general
rule to match the majority, e.g.\ separating \acronym{SMTP} delivery to
specific sites from \acronym{SMTP} delivery to the rest of the world.  The
algorithm provides a mechanism for ordering overlapping rules: the priority
field in each rule (defaults to zero).  Negative priorities may be useful
for catchall rules.

Detecting overlapping rules is difficult, but the following approaches may
be helpful:

    \item Sort by regex and visually inspect the list, e.g.\ with
        \acronym{SQL} similar to: \textbf{select regex from rules order by
        regex;}

    \item Compare the results of parsing using sorted, shuffled, and
        reversed rules (described in \sectionref{rule ordering for
        efficiency}).  Parse several log files using optimal ordering, then
        dump a textual representation of the rules, connections, and
        results tables.  Repeat with shuffled and reversed ordering,
        starting with a fresh database.  If there are no overlapping rules
        the tables from each run will be identical; differences indicate
        overlapping rules.  The rules that overlap can be determined by
        examining the differences in the tables: each result contains a
        reference to the rule which created it, if the references differ
        between runs the two rules referenced in the differing results
        overlap.  Unfortunately this method cannot prove the absence of
        overlapping rules; it can detect overlapping rules, but only if
        there are log lines in the log files that are matched by more than
        one rule.

>>>>

Rules categorise inputs, specifying both the regex to match against each
input and the action to invoke when the match is successful.  Parsing new
inputs is generally achieved by creating a new rule that pairs an existing
action with a new regex.  Decoupling the rules from the actions and
framework enables other rule management approaches to be used, e.g.\
instead of manually adding new rules, machine learning techniques could be
used to automatically generate new rules.  If this approach was taken the
choice of machine learning technique would be constrained by the size of
typical data sets (see \sectionref{Results} XXX CHECK THIS REFERENCE WHEN
THE RESULTS SECTION IS FINISHED).  Techniques requiring the full data set
when training would be impractical; Instance Based
Learning~\cite{instance-based-learning} techniques that automatically
determine which inputs from the training set are valuable and which inputs
can be discarded might reduce the data required to a manageable size.  A
parser might also dynamically create new rules in response to certain
inputs, e.g.\ diagnostic messages indicating the program which produced the
input being parsed had read a new configuration file.  These avenues of
research and development has not been pursued by the author, but could
easily be undertaken independently.

The architecture does not try to detect overlapping rules: that
responsibility is left to the author of the rules.  Unintentionally
overlapping rules lead to inconsistent parsing and data extraction because
the first matching rule wins, and the order in which rules are tried
against each input might change between parser invocations.  Overlapping
rules are frequently a requirement, allowing a more specific rule to match
some inputs and a more general rule to match the remainder, e.g.\
separating \acronym{SMTP} delivery to specific sites from \acronym{SMTP}
delivery to the rest of the world.  Allowing overlapping rules simplifies
both the general rule and the more specific rule; additionally rules from
different sources can be combined with a minimum of prior cooperation or
modification required.  Overlapping rules should have a priority attribute
to specify their relative ordering; negative priorities may be useful for
catchall rules.

Decoupling the rules from the actions allows external tools to be written
to detect overlapping rules.  Traditional regexes are equivalent in
computational power to \acronym{FA} and can be converted to \acronym{FA},
so regex overlap can be detected by finding a non-empty intersection of two
\acronym{FA}\@.  The standard equation for \acronym{FA} intersection (given
for example in~\cite{intersection-of-NFA-using-Z}) is: $FA1 \cap{} FA2 =
\overline{(\overline{FA1} \cup{} \overline{FA2})}$, which has considerable
computation complexity.  Perl 5.10 regexes are more powerful than
traditional regexes: it is possible to match correctly balanced brackets
nested to an arbitrary depth, e.g.\
\verb!/^[^<>]*(<(?:(?>[^<>]+)|(?1))*>)[^<>]*$/! matches
\verb!z<123<pq<>rs>j<r>ml>s!.  Perl 5.10 regexes can maintain an arbitrary
state stack and are thus equivalent in computational power to \acronym{PDA}
or \acronym{CFL}, so detecting overlap may require calculating the
intersection of two \acronym{PDA} or \acronym{CFL}s.  The intersection of
two \acronym{CFL}s is not closed, i.e.\ the resulting language cannot
always be parsed by a \acronym{CFL}, so intersection may be intractable in
some cases e.g.:
$a^{*}b^{n}c^{n}~\cap~a^{n}b^{n}c^{*}~\rightarrow~a^{n}b^{n}c^{n}$.
Detecting overlap amongst $n$ regexes requires calculating $n(n-1)/2$
intersections, resulting in $O(n^2x)$ complexity, where $O(x)$ is the
complexity of calculating intersection.  This is certainly not a task to be
performed every time the parser is used: detecting overlap amongst the
Postfix log parser's \numberOFrules{} rules would require calculating
\numberOFruleINTERSECTIONS{} intersections.

The framework requires each rule to have \texttt{action} and \texttt{regex}
attributes; each implementation is free to add any additional attributes it
requires.

It is possible to define pathological regexes which fall into two main
categories: regexes that match every input, and regexes that consume
excessive amounts of CPU time during matching.  Defining a regex to match
all inputs is trivial: \verb!/^/! matches the start of every input.
Usually excessive CPU time is consumed when a regex with a lot of
alteration and variable quantifiers fails to match, but successful
matching is generally quite fast (see~\cite{mastering-regular-expressions}
for in-depth discussion).

XXX RULES CAN HAVE CONDITIONS ATTACHED

@@

Rule have certain characteristics that may help in understanding the
parser architecture:

XXX IMPROVE THIS SECTION\@; SHOULD IT BE A SUBSECTION\@?

@@

    \item Rules are annotated with the name of a Postfix program, and will
        only be used when parsing log lines produced by that
        program.

@@

\subsection{Architecture Characteristics}

\label{Architecture characteristics}

\begin{description}

    \item [Matching rules against inputs is simple:]  The first matching
        rule determines the action that will be invoked: there is no
        backtracking to try alternate rules, no attempt is made to pick a
        \textit{best\/} rule.

    \item [Line oriented:]  The architecture is line oriented at present:
        there is no facility for rules to consume more input or push unused
        input back onto the input stream.  This was not a deliberate design
        decision, rather a consequence of the line oriented nature of
        Postfix log files; more flexible approaches could be pursued.

    \item [Context-free rules:]  Rules can not take into account past or
        future inputs.  In context-free grammar terms the parser rules
        could be described as:
        \newline{}$\text{\textless{}input\textgreater{}} \mapsto
        \text{rule-1} | \text{rule-2} | \text{rule-3} | \dots |
        \text{rule-n}$.

    \item [Context-aware actions:] Actions can consult the results (or lack
        of results) of previous actions during execution, providing some
        context sensitivity.

    \item [Cascaded parsing:] Actions can return a modified input to be
        parsed as if read from the input stream, allowing for a simplified
        version of cascaded parsing~\cite{cascaded-parsing}.

    \item [Transduction:]  The architecture can be thought of as
        implementing transduction: it takes data in one form (log files)
        and transforms it to another form (a database); other formats may
        be more suitable for other implementations.

    \item [Similarity to \acronym{NLP}:] \hfill{} \newline{}
        Unlike traditional parsers such as those used when compiling a
        programming language, this architecture does not require a fixed
        grammar specification that inputs must adhere to.  The architecture
        is capable of dealing with interleaved inputs, out of order inputs,
        and ambiguous inputs where heuristics must be applied --- all have
        arisen and been successfully accommodated in the Postfix log
        parser.

\end{description}

### 2009-02-22 doc/logparser-implementation.tex
The beginning of the architecture rules section.

\subsection{Using a Database as an Application Programming Interface}

@@

>>>>

\subsection{Using A Database To Provide An Application Programming Interface}

@@

\subsection{Overlapping Rules}

\label{overlapping rules in architecture}

XXX IS THERE ANY PREVIOUS RESEARCH IN THIS AREA\@?

XXX DFA COMPARISON\@.

The parser does not try to detect overlapping rules; that responsibility is
left to the author of the rules.  Unintentionally overlapping rules lead to
inconsistent parsing and data extraction because the order in which rules
are tried against each line may change between log files, and the first
matching rule wins.  Overlapping rules are frequently a requirement,
allowing a more specific rule to match some log lines and a more general
rule to match the majority, e.g.\ separating \acronym{SMTP} delivery to
specific sites from \acronym{SMTP} delivery to the rest of the world.  The
algorithm provides a mechanism for ordering overlapping rules: the priority
field in each rule (defaults to zero).  Negative priorities may be useful
for catchall rules.

Detecting overlapping rules is difficult, but the following approaches may
be helpful:

    \item Sort by regex and visually inspect the list, e.g.\ with
        \acronym{SQL} similar to: \textbf{select regex from rules order by
        regex;}

    \item Compare the results of parsing using sorted, shuffled, and
        reversed rules (described in \sectionref{rule ordering for
        efficiency}).  Parse several log files using optimal ordering, then
        dump a textual representation of the rules, connections, and
        results tables.  Repeat with shuffled and reversed ordering,
        starting with a fresh database.  If there are no overlapping rules
        the tables from each run will be identical; differences indicate
        overlapping rules.  The rules that overlap can be determined by
        examining the differences in the tables: each result contains a
        reference to the rule which created it, if the references differ
        between runs the two rules referenced in the differing results
        overlap.  Unfortunately this method cannot prove the absence of
        overlapping rules; it can detect overlapping rules, but only if
        there are log lines in the log files that are matched by more than
        one rule.
